---
title: "Brownian Motion with Ito and Stratonovich"
author: "Graeme Keleher"
date: "June 7, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Install Packages
The focus of this demonstration is the "Sim.DiffPrc" package which is designed for simulating diffusion processes. Click the link for full documentation:

https://cran.r-project.org/web/packages/Sim.DiffProc/Sim.DiffProc.pdf

ggplot2 is a package for graphing, tidyr and dplyr are for data manipulation.

```{r message=FALSE, warning=FALSE}
library(Sim.DiffProc)
library(ggplot2)
library(dplyr)
library(tidyr)
```

#Ito Method Realizations of Basic Browian Motion 
For basic Brownian motion with the martingale property there is no drift. My choice of 1 as the diffusion coefficent is arbitrary.

```{r}
#Specify the drift and diffusion coefficents
f = expression(0)
g = expression(1)

#Specify desired number of realizations
number = 50

#Returns an 'snssde1d' object containing realizations of the specified diffusion process 
Realizations = snssde1d(drift=f,diffusion=g,M=number, type = "ito")

#Returns the final positions of the realizations (i.e. at position t = 1)
End_Position = as.data.frame(rsde1d(Realizations, at = 1))
colnames(End_Position) = c("Position")

#Extract the time series' from the 'snssde1d' object and transform to a data frame 
#to make plotting with ggplot2 possible
TS =  as.data.frame(Realizations$X)
```

#Plots For Ito Realizations
```{r, echo = FALSE, message=FALSE, warning=FALSE}

graph_TS =  mutate(TS,id = row_number()) %>%
  gather(Realization,value, -id)

graph_TS$id = graph_TS$id/1000
  
ggplot(graph_TS,aes(x = id, y = value, color = Realization)) + geom_line() +
  labs(title = "Ito Method Sample Paths", x = "Time", y = "Displacement")+ theme(legend.position="none")

ggplot(End_Position, aes(Position)) + geom_histogram() +
  labs(title = "Histogram of Ito Method Realization End Positions", x = "Position at Time T = 1", y = "")

```

#Stratonovich Method Realizations of Basic Browian Motion
I am replicating the above example with the only change being the method. 

```{r}
#Specify the drift and diffusion coefficents
f = expression(0)
g = expression(1)

#Specify desired number of realizations
number = 50

#Returns an 'snssde1d' object containing realizations of the specified diffusion process 
#The type has changed from ito to Stratonovich
Realizations <- snssde1d(drift=f,diffusion=g,M=number, type = "str")

#Returns the final positions of the realizations (i.e. at position t = 1)
End_Position <- as.data.frame(rsde1d(Realizations, at = 1))
colnames(End_Position) = c("Position")

#Extract the time series' from the 'snssde1d' object and transform to a data frame 
#to make plotting with ggplot2 possible
TS =  as.data.frame(Realizations$X)
```

#Plots For Stratonovich Method Realizations
```{r, echo= FALSE, message=FALSE, warning=FALSE}
graph_TS =  mutate(TS,id = row_number()) %>%
  gather(Realization,value, -id)

graph_TS$id = graph_TS$id/1000
  
ggplot(graph_TS,aes(x = id, y = value, color = Realization)) + geom_line() +
  labs(title = "Stratonovich Method Sample Paths", x = "Time", y = "Displacement")+ theme(legend.position="none")

ggplot(End_Position, aes(Position)) + geom_histogram() +
  labs(title = "Histogram of Stratonovich Method Realization End Positions", x = "Position at Time T = 1", y = "")

```

#Ito Method Realizations of Geometric Browian Motion
Geometric Brownian motion has drift and diffusion coefficients that are functions of position. Because of this, the process cannot start at zero if we want non-trivial results. 

```{r}
#Specify the drift and diffusion coefficents
f = expression(3*x)
g = expression(.5*x)

#Specify desired number of realizations
number = 50

#Returns an 'snssde1d' object containing realizations of the specified diffusion process.
#Note that the x0 value is no longer 0.
Realizations = snssde1d(drift=f,diffusion=g,M=number, type = "ito", x0 = 1)

#Returns the final positions of the realizations (i.e. at position t = 1)
End_Position = as.data.frame(rsde1d(Realizations, at = 1))
colnames(End_Position) = c("Position")

#Extract the time series' from the 'snssde1d' object and transform to a data frame 
#to make plotting with ggplot2 possible
TS =  as.data.frame(Realizations$X)
```

#Plots For Ito Realizations Of Geometric Brownian Motion
```{r ,echo= FALSE, message=FALSE, warning=FALSE}

graph_TS =  mutate(TS,id = row_number()) %>%
  gather(Realization,value, -id)

graph_TS$id = graph_TS$id/1000
  
ggplot(graph_TS,aes(x = id, y = value, color = Realization)) + geom_line() +
  labs(title = "Sample Paths of Geometric Brownian Motion", x = "Time", y = "Displacement")+ theme(legend.position="none")

ggplot(End_Position, aes(Position)) + geom_histogram() +
  labs(title = "Histogram of Geometric Brownian Motion End Positions", x = "Position at Time T = 1", y = "")

```
